---
title: "Assignment8_Seshadri"
author: "Sri Seshadri"
date: "8/12/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=70),tidy=TRUE)
library(magrittr)
```
## Assignment
In this assignment we will learn how to perform an exploratory data analysis for a clustering problem, fit a hierarchical cluster analysis, fit a k-means cluster analysis, how to integrate principal components analysis and cluster analysis, how to use cluster analysis as a predictive model, and how to make a variety of R graphics applicable to cluster analysis and multivariate analysis in general. 

## Part 1- The data
Let’s begin by reading in the data.


```{r}
my.data <- read.csv('European_Employment.csv', header = T)
levels(my.data$Group)
pairs(my.data[,-2], col = my.data$Group)

```

Appears that FIN vs SER and MAN vs FIN are interesting

## 2D plots



```{r pressure, echo=FALSE}
library(ggplot2)
ggplot(data = my.data, mapping = aes(x = SER, y = FIN, color = Group, label = Country)) + geom_point() + geom_text(show.legend = F) + theme_bw() + xlab("Service") + ylab("Finance")

ggplot(data = my.data, mapping = aes(x = MAN, y = FIN, color = Group, label = Country)) + geom_point() + geom_text(show.legend = F) + theme_bw() + xlab("Manufacturing") + ylab("Finance")
```

## PCA
Each row is desogned to add to 100%. So if the data is standardized and scaled, we may lose the property. Hence PCA will be done without standardization.
```{r}
pca <- princomp(my.data[,c(-1,-2)], cor = F)
summary(pca)
my.data$pc.1 <- pca$score[,1]
my.data$pc.2 <- pca$score[,2]
ggplot(data = my.data,mapping = aes(x = pc.1, y = pc.2, color = Group, label = Country)) + geom_point() + geom_text(size = 3,show.legend = F) + theme_light()

```


# Hirearchical Clustering analysis

Let the drop the group - "other". And what does the cluster look like when we attempt to reproduce the FIN vs Ser

```{r}
my.data <- dplyr::as_tibble(my.data)
subset <- my.data %>% 
  dplyr::filter(Group != "Other") 
fin.ser <- hclust(dist(subset[,c('FIN','SER')]),method = "complete")
plot(fin.ser,labels = subset$Country)

```

```{r}
class <- cutree(fin.ser,k = 3)
class2 <- cutree(fin.ser,h = 12) # class 1 & 2 are the same
subset$clust3 <- class
tab <- table(subset$Group,subset$clust3)
tab <- tab[1:3,]
# t(tab) * (1/apply(tab,FUN = sum, MARGIN = 2))
# tab

round(t(tab) / colSums(tab),2)
class6 <- cutree(fin.ser,k=6)
subset$clust6 <- class6
tab2 <- table(subset$Group,subset$clust6)
tab2 <- tab2[1:3,]
# t(tab) * (1/apply(tab,FUN = sum, MARGIN = 2))
# tab
round(t(tab2) / colSums(tab2),2)


```

## Lets try a PCA again

```{r}
pca.out <- princomp(subset[,3:11],cor = F)
summary(pca.out)
subset$pc.out.1 <- pca.out$scores[,1]
subset$pc.out.2 <- pca.out$scores[,2]
pca.hclust <- hclust(dist(subset[,c('pc.out.1','pc.out.2')]), method = 'complete')
plot(pca.hclust, labels = subset$Country)

tab3 <- table(subset$Group,cutree(pca.hclust,k = 3))
round(t(tab3)/colSums(tab3),2)
tab4 <- table(subset$Group,cutree(pca.hclust,k = 6))
round(t(tab4)/colSums(tab4),2)
```

### K means clustering

```{r}

finser <- kmeans(x = subset[,c('FIN','SER')], centers = 3)
subset$kmeans3 <- finser$cluster
centers <- as.data.frame(finser$centers)
ggplot() + geom_point(data = subset, mapping = aes(y = FIN, x = SER, color = Group)) + geom_text(data = subset, mapping = aes(y = FIN, x = SER, color = Group, label = Country ),show.legend = F,size = 3,nudge_x = 0.3,nudge_y = 0.01) +
  geom_point(data = centers , mapping = aes(y = FIN, x = SER),shape = 23, fill = 'red', size = 5) + geom_text(data = centers , mapping = aes(y = FIN, x = SER, label = rownames(centers)),size = 3) + theme_bw()
tab5 <- table(subset$Group, subset$kmeans3)
t(tab5)/colSums(tab5)


finser <- kmeans(x = subset[,c('FIN','SER')], centers = 6)
subset$kmeans6 <- finser$cluster
centers <- as.data.frame(finser$centers)
ggplot() + geom_point(data = subset, mapping = aes(y = FIN, x = SER, color = Group)) + geom_text(data = subset, mapping = aes(y = FIN, x = SER, color = Group, label = Country ),show.legend = F,size = 3,nudge_x = 0.3,nudge_y = 0.01) +
  geom_point(data = centers , mapping = aes(y = FIN, x = SER),shape = 23, fill = 'red', size = 5) + geom_text(data = centers , mapping = aes(y = FIN, x = SER, label = rownames(centers)),size = 3) + theme_bw()
tab6 <- table(subset$Group, subset$kmeans6)
t(tab6)/colSums(tab6)

```



## Hmmmm.... Let's do this on the principal component scores

```{r}
finser <- kmeans(x = subset[,c('pc.out.1','pc.out.2')], centers = 3)
subset$kmeans3 <- finser$cluster
centers <- as.data.frame(finser$centers)
ggplot() + geom_point(data = subset, mapping = aes(y = pc.out.2, x = pc.out.1, color = Group)) + geom_text(data = subset, mapping = aes(y = pc.out.2, x = pc.out.1, color = Group, label = Country ),show.legend = F,size = 3,nudge_x = 0.3,nudge_y = 0.01) +
  geom_point(data = centers , mapping = aes(y = pc.out.2, x = pc.out.1),shape = 23, fill = 'red', size = 5) + geom_text(data = centers , mapping = aes(y = pc.out.2, x = pc.out.1, label = rownames(centers)),size = 3) + theme_bw()
tab5 <- table(subset$Group, subset$kmeans3)
t(tab5)/colSums(tab5)

finser <- kmeans(x = subset[,c('pc.out.1','pc.out.2')], centers = 6)
subset$kmeans3 <- finser$cluster
centers <- as.data.frame(finser$centers)
ggplot() + geom_point(data = subset, mapping = aes(y = pc.out.2, x = pc.out.1, color = Group)) + geom_text(data = subset, mapping = aes(y = pc.out.2, x = pc.out.1, color = Group, label = Country ),show.legend = F,size = 3,nudge_x = 0.3,nudge_y = 0.01) +
  geom_point(data = centers , mapping = aes(y = pc.out.2, x = pc.out.1),shape = 23, fill = 'red', size = 5) + geom_text(data = centers , mapping = aes(y = pc.out.2, x = pc.out.1, label = rownames(centers)),size = 3) + theme_bw()
tab5 <- table(subset$Group, subset$kmeans3)
t(tab5)/colSums(tab5)
sum(tab5[1:3,])
```


```{r}
Accuracy <- function(cluster,data = subset) {
  # lets get the hierarchical clustering accuracy
  hclust <- hclust(dist(data[,c("AGR","MIN","MAN","PS","CON","SER","FIN","SPS","TC" )]), method = "complete")
  data$hiercluster <- cutree(hclust,k = cluster)
  tab <- table(data$Group,data$hiercluster)
  hier.Accuracy <- ifelse(cluster == 1,max(tab[1:3,])/sum(tab[1:3,]),sum(apply(tab[1:3,],MARGIN = 2,FUN = max))/sum(colSums(tab[1:3,])))
  
  # k means 
  kclust <- kmeans(data[,c("AGR","MIN","MAN","PS","CON","SER","FIN","SPS","TC" )],centers = cluster)
  data$kclust <- kclust$cluster
  tab2 <- table(data$Group, data$kclust)
  kmean.Accuracy <- ifelse(cluster == 1,max(tab[1:3,])/sum(tab[1:3,]),sum(apply(tab2[1:3,],MARGIN = 2,FUN = max))/sum(colSums(tab2[1:3,])))
  
  return(data.frame(cluster = cluster,hier.Accuracy = hier.Accuracy, kmean.Accuracy = kmean.Accuracy))
}

AccuracyResults <-  purrr::map_df(.x = 1:6,.f = Accuracy)
ggplot(data = AccuracyResults, mapping = aes(x = cluster)) + geom_line(aes(y = hier.Accuracy, color = 'red'), show.legend = F ) + geom_point(aes(y = hier.Accuracy, color = 'red'),show.legend = F ) + geom_line(aes(y = kmean.Accuracy, color = 'blue'),show.legend = F) + geom_point(aes(y = kmean.Accuracy, color = 'blue') ,show.legend = F ) + theme_bw() + xlab("# Clusters") + ylab("Accuracy") 
```






























