---
title: "ClusterAnalysis"
author: "Sri Seshadri"
date: "8/9/2017"
output:
  html_document: default
  always_allow_html: yes
---

```{r setup, include=FALSE, cutoff.width = 70}
knitr::opts_chunk$set(echo = TRUE)
measure <- readRDS('measure.RDS')
library(plotly)
library(graphics)
```

## EDA

Since we know what the data is, let's try a 3d plot... colored by gender

```{r measure, warning=FALSE, message=F}
p <- plotly::plot_ly(data = measure, x = ~chest, y = ~waist, z = ~hips , type = 'scatter3d',mode = 'markers',color = ~gender)
# Sys.setenv("plotly_username"="srivathsesh")
# Sys.setenv("plotly_api_key"="0Sr49roqMwTtuQl3JZ9G")
# api_create(p, filename = "r-clusterAnalysis")
# knitr::include_graphics('https://plot.ly/~srivathsesh/5.embed')
p
```

## Great, now what??

Let's see if we can create the visual clustering mathematically...

Lets start with the eucledian distance. Lets manually create the euclidean distance between first row and the second row
$$\sqrt{(measure[1,1] -  measure[2,1])^2 + ( measure[1,2] - measure[2,2])^2 + (measure[1,3] - measure[2,3])^2}$$
`r sqrt((measure[1,1] -  measure[2,1])^2 + ( measure[1,2] - measure[2,2])^2 + (measure[1,3] - measure[2,3])^2)`



```{r, message=F,warning=F }
# create a matrix of the euclidean distances
dm <- dist(measure[1:3], method = "euclidean" )
single <- hclust(dm,method = "single")
plot(single,main = "Single")
Complete <- hclust(dm,method = "complete")
plot(Complete,main =  "Complete")
Average <- hclust(dm,method = "average")
#plot(as.dendrogram(Average),main = "Average",horiz = T)
plot(Average,main = "Average")
```

### How do they compare to the principle components ?

```{r}
pc <- princomp(measure[,1:3],cor = T)
summary(pc)
pc2 <- princomp(dm, cor = T)
summary(pc2)
measure$pc1 <- pc$scores[,1]
measure$pc2 <- pc$scores[,2]
measure$dmpc1 <- pc2$scores[,1]
measure$dmpc2 <- pc2$scores[,2]
measure$single <- cutree(single,h = 3.6)
measure$complete <- cutree(Complete, h = 10)
measure$avg <- cutree(Average, h = 7.9)
plot_ly(data = measure,x = ~pc1, y = ~pc2, color = ~gender, type = 'scatter', mode = 'markers',colors = "Set1") %>% 
  add_text(text = measure$avg, textposition = "top right",showlegend = F) %>% 
  layout(title = 'PCA using raw data and class labels based on Average')

plot_ly(data = measure,x = ~dmpc1, y = ~dmpc2, color = ~gender, type = 'scatter', mode = 'markers',colors = "Set1") %>% 
  add_text(text = measure$avg, textposition = "top right",showlegend = F) %>% 
  layout(title = 'PCA using Eucleadian distances and class labels based on Average')
```
## K-Means Clustering - Textbook example

```{r}
crime <- readRDS(file = 'crime.rds')

symbol <- rep(20,51)
symbol[which(crime$Murder > 15)] <- 3
pairs(crime,pch = symbol)
```
The "+" symbol indicates the record for MD... did you see that its an outlier. Lets leave that out data point

```{r}
crime_subset <- dplyr::filter(crime, rownames(crime) != 'DC')
rownames(crime_subset) <- rownames(crime[which(rownames(crime) != 'DC'),])
sapply(crime_subset,'var')
```

See the variance is different for the different variables? It necessary standardize the variables... Not necessarily scale it. The scaling doesn't matter as we are interested in the distances within each axis (variable... see the euclidean formula above.) 

```{r}
# Get the ranges for each column
rge <- sapply(crime_subset, function(x) diff(range(x)))
crime_s <- sweep(crime_subset,2,rge,FUN = "/")
rownames(crime_s) <- rownames(crime_subset)
n <- nrow(crime_s)
OverallSS <- (n-1) * sum(sapply(crime_s,var))

kmeanscree <- function(data,center) {
  withinss <- purrr::map(.x = center, .f = function(x) kmeans(data,center =x)$tot.withinss)
  withinss <- c((nrow(data) - 1) * sum(sapply(data,var)),withinss)
  plot(x = c(1,center), y = unlist(withinss),xlab = 'Number of clusters',ylab = 'Within cluster Sum of Squares',type = 'b' )
}
kmeanscree(data = crime_s,center = 2:6)

```

2 clusters form the elbow of the scree plot. We'll use the 2 clusters for the kmean clustering.

```{r}
kclust <- kmeans(x = crime_s, centers = 2)
# converting back to non standardized units
kclust$center*rge
# PCA
pca.crime <- princomp(x = crime_s,cor = F)
summary(pca.crime)
# populate df for plotting
df.pca <- data.frame(pc1 = pca.crime$scores[,1], pc2 = pca.crime$scores[,2],clusters = as.factor(kclust$cluster),Index = rownames(crime_s))
ggplot(df.pca,mapping = aes(x = pc1, y = pc2, color = clusters,label = Index)) + geom_point() + geom_text(size = 3) + theme_bw()
```


